---
title: "RMED901 Group 3 report"
Authors: "Amalie Austgulen, André Høberg, Ariell Thale Toftesund Longva, Ghazal Lessan Toussi"
output: html_document
date: "2022-09-15"
---
This is the report for group 3 in RMED901.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Reading and loading the data**  

We load the necessary packages first.
```{r, echo = FALSE}
library(tidyverse)
library(here)
```

We read the non tidy version of the data, that we will try to tidy later.
```{rm echo = FALSE}
data_nontidy <- read_delim(here("data", "copy_exam_nontidy.txt"))
```

**Exploring the data**  

Before tidying, we wished to explore the data. We did this by using functions such as head, tail, summary and glimpse.  
Here we only show the results of summary and glimpse, since these gave us the most information about the dataset.
```{r, echo = FALSE}
head(data_nontidy)
tail(data_nontidy)
```

```{r}
summary(data_nontidy)
```
```{r}
glimpse(data_nontidy)
```
```{r}
skimr::skim(data_nontidy)
```

There are 15 columns and 34 048 rows in this dataset. 

When looking at the data, we noticed multiple aspects that we wished to change later, such as:  

- Time.measurement should be two variables: col_rec_that and rec_ver_tat
- Gender.age was one column, when it should be two  
- Age, pan_day, drive_thru_ind and ct_result were all of the type character, when they should be numeric.  

We also noticed that there are 12 344 unique values for subject, which may indicate that there are variables that lead to multiple observations.  

In the skim, we saw that there is 15527 missing values for the column payor_group, 15506 for patient_class and 452 for ct_result. This we explored further with the use of gg_miss_var.

```{r}
naniar::gg_miss_var(data_nontidy)
```





